---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

This assignment aims for the analysis of activity monitoring of steps taken in 5-minute interval. Next, the process of analysis is described and R code is presented.

## Loading and preprocessing the data

The data is loaded using read.csv function. Afterthat, the date is converted from character format into date format.

```{r echo=TRUE}
activity <- read.csv("activity.csv", stringsAsFactors=F)
activity$date2 <- strptime (activity$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

An histogram of the total number of steps taken each day is created.

```{r echo=TRUE}
attach(activity)
aggdata <-aggregate(activity$steps, by=list(date), FUN=sum, na.rm=TRUE)
print(aggdata)
detach(activity)

library(ggplot2)
plot <- ggplot(aggdata, aes(x=x))
plot <- plot +  geom_histogram(aes(y=..density..), binwidth = 500, colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")  # Overlay with transparent density plot
print(plot)
```

The mean and the median of the total number of steps are calculated. Then, mean (red line) and the median (blue line) are plotted in the histogram created abode.

```{r echo=TRUE}
mean <- mean(aggdata$x)
median <- median(aggdata$x)
plot <- plot + geom_vline(aes(xintercept=mean), color="red", linetype="dashed", size=1) + geom_vline(aes(xintercept=median), color="blue", linetype="dashed", size=1)
print(plot)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
